Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-01-24 thesis.qmd
lilly madau explanation

@@

SFR can be estimated using various photometric or spectroscopic methods based on the luminosity of at least one spectral band or the intensity of a spectral line. Different luminosities and intensities trace distinct emission mechanisms, offering insights into a galaxy’s radiation sources. Below are common methods:[^2]

[^2]: @calzettiStarFormationRate2012, @mushotzkyASTR620GalaxiesFall2017

@@

## Delayed-$τ$ Model

# Lilly-Madau Plot

@@

![](figure/lily-madau.png){fig-align="center" width="450"}

# Delayed-$\tau$ model

More details about the model and how it clashes with the Lilly-Madau plot

>>>>

It is also important to define the Star Formation Density (SFRD):

$$
\text{SFRD} = \frac{1}{V}\sum^N_{i=1}\text{SFR}_{o,i},
$$ where $V$ is the comoving volume [^2] of a region of the universe, $N$ is the number of galaxies of the regions and SFR$_{0,i}$ is the current SFR of each galaxy $i$.

[^2]: Comoving volume factors out the expansion of the universe, it remains a constant measure of “space” over time, making it possible to compare star formation at different epochs

For the Local Cosmological Volume, which we will focus on, the volume is the sphere with its center at the Milky Way and a radius of 11 Mpc $V = \frac{4}{3}\pi(11\, \text{Mpc})^3\approx 5575.3\, \text{Mpc}^3$.

@@

SFR can be estimated using various photometric or spectroscopic methods based on the luminosity of at least one spectral band or the intensity of a spectral line. Different luminosities and intensities trace distinct emission mechanisms, offering insights into a galaxy’s radiation sources. Below are common methods:[^3]

[^3]: @calzettiStarFormationRate2012, @mushotzkyASTR620GalaxiesFall2017

@@

# Lilly-Madau Plot and Delayed-$\tau$ model conflicts

@@

![The history of cosmic star formation as shown in @madauCosmicStarFormation2014](figure/lily-madau.png){fig-align="center" width="450"}

Using observational data from UV and IR traces for the SFR for galaxies of different redshifts ($z$) @madauCosmicStarFormation2014 where able to calculate the cosmic SFRD, at different ages of the universe, and reveals three periods of the universe.

-   **Rising Era** (SFRD$\propto(1+z)^{-2.9},\ 3\lesssim z \lesssim 8$): The SFRD is relatively low at very high redshifts (e.g., $z>6$) and then rises as we move closer to $z \sim 2$. This indicates that early galaxies were ramping up their star-forming activity.

-   **Peak Star Formation** ($z \sim 2$): Often dubbed “cosmic noon,” this epoch exhibits the highest SFRD in cosmic history when the Universe was $\sim 3.5$ Gyr old. Galaxies are vigorously converting gas into stars, and many of the most massive galaxy systems today formed much of their stellar content during this phase.

-   **Decline to Present Day **(SFRD$\propto (1+z)^{2.7},z\lesssim 2$): After $z \sim 2$, the global SFRD drops by about an order of magnitude towards the present day ($z=0$).

The best-fit function of the comoving SFRD($z$):

$$
SFRD(z) = 0.015 \, \frac{(1+z)^{2.7}}{1 + \left(\frac{1+z}{2.9}\right)^{5.6}}
\quad \left[M_\odot \,\mathrm{yr}^{-1}\,\mathrm{Mpc}^{-3}\right],
$$

## Delayed-$\tau$ model
